---
title: "Get untransformed fluorescence values"
author: "Julie Chuong"
date: "Sep 13 2022"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
Get untransformed fluorescence values. 
Additionally try to see if we can transform them such in a way that the 0,1,2 copy controls
are in a linear scale. 


Install CytoExplorer package and requirements (can be skipped if already installed)
```{r}
library(BiocManager)
install.packages("cytolib", "flowCore", "flowWorkspace", "openCyto")
```

Install CytoExploreR from GitHub:
```{r}
library(devtools)
devtools::install_github("DillonHammill/CytoExploreR")
```

Load required packages
```{r}
library(CytoExploreR)
library(tidyverse)
library(ggridges)
library(docstring)
```

Set working directory and get list of subdirectories containing FCS files
```{r setup}
#setwd("/Volumes/GoogleDrive/My Drive/greshamlab/Molecular Determinants of CNV Evolution Dynamics/Summer 2021 Group LTEE/FCS files") #Julie's WD
knitr::opts_knit$set(root.dir = "/Volumes/GoogleDrive/My Drive/greshamlab/projects/EE_GAP1_ArchMuts_Summer2021/data/Summer_LTEE_2021_FCS_files")  #Julie's WD
```

select the FSC file folders in your directory
```{r}
getwd()
folders = list.dirs()[c(7:30)]
```

# Choose a name to be used for all output files including the gating template and associated flow data and graphs.
```{r}
version_name = "raw_flow"
```

STEP 1: Generate experiment details file from folder and FCS file names
Experiment details file is a .csv file that contains the list of .fcs files in the directory and the associated metadata for each sample
Author: Grace

```{r}
make_exp_details = function(folder_name, samplesheet) {
  pref = folder_name %>% str_extract("([0-9])+_EE_GAP1_ArchMuts_2021")
  generation = folder_name %>% str_extract("[g]\\d+") %>% str_remove("g")

  files = as_tibble(list.files(paste0(folder_name))) %>%
    separate(value, into = c("well", "samp"), sep = " ", remove = F) %>%
    mutate(well = str_extract(well, "([A-Z])([0-9]){1,2}$")) %>%
    mutate(samp = str_remove(samp, ".fcs")) %>%
    mutate(sample = case_when(str_detect(value, "Unstained") ~ "ctrl0",
                              str_detect(value, "DGY500") ~ "ctrl1",
                              str_detect(value, "DGY1315") ~ "ctrl2",
                              TRUE ~ samp)) %>%
    select(value,sample) %>%
    rename(name = value) %>%
    filter(!is.na(sample))

  all = files %>%
    left_join(read_csv(paste0("./",samplesheet)), by = c("sample" = "Sample name")) %>%
    mutate(generation = as.numeric(generation))

  write_csv(all, file = paste0(folder_name,"/",pref,"_experiment_details.csv"))

}

#needs to be run once
map(folders, make_exp_details, samplesheet = "EE_GAP1_ArchMuts_2021.csv")
```


#STEP 2: Read in all files in a directory and rename the channels.
#A directory contains an FCS file for each population.
#Results in 1 timepoint gating set containing all .fcs files, associated experiment details, and marker details
#Author: Julie

# here we will load in 1 directory that contains 3 timepoints worth of data to load in.
# cyto_setup() does not permit loading in more than 1 directory, so I had to create a directory with the data files of interest.
# these data will guide us on drawing gates.
```{r}
exp_details_path = list.files(path = paste0(folders[1]), pattern = "_experiment_details.csv", full.names = T)
exp_details_path
```

```{r}
timepoint_gating_set <- cyto_setup(path = paste0(folders[1]), restrict=TRUE, select="fcs", details=F) #edit Markers on Viewer pane, Save & Close
```

Use flowWorkspace::pData to annotate the experiment details file associated with the gating set
```{r}
experiment_details <- read_csv(exp_details_path) #import experiment-details.csv

for(i in 1:length(names(experiment_details))){
  flowWorkspace::pData(timepoint_gating_set)[names(experiment_details[i])]<-experiment_details[i]
}

cyto_details(timepoint_gating_set)
```

Rename the experiment-markers.csv file. Need to do once.
```{r}
file.rename(dir(pattern = "Experiment-Markers.csv"),"EE_GAP1_ArchMuts_2021-Experiment-Markers.csv")
```


We want to skip the transformation and just output the raw fluorescence values

Generate single cell data tables and normalized fluorescence

Extract B2-A and FSC-A values
```{r}
timepoint_raw_list <- cyto_extract(timepoint_gating_set, raw = T, channels = c("FSC-A", "B2-A"))
```

```{r}
norm_raw=map_df(timepoint_raw_list, ~as.data.frame(.x), .id="name") %>% #convert to df, put list name in new column
  mutate(name = as.factor(name)) %>% #convert `name` to factor
  left_join(experiment_details %>% #join by name column to add metadata
  mutate(generation = as.factor(unique(experiment_details$generation)))) %>%
  mutate(B2A_FSC = `B2-A`/`FSC-A`) #compute normalized fluor
 # write_csv(paste0(version_name,"_SingleCellDistributions_",prefix,".csv"))

norm_raw %>%
  group_by(Description) %>%
  summarize(median(B2A_FSC))

hist(norm_raw$`B2-A`)
hist(norm_raw$`FSC-A`)
```
### See if different transformation can make the 0, 1 and 2 copy controls scale linearly. 
transform the data

looks useful if I want to choose different transformation: https://dillonhammill.github.io/CytoExploreR/articles/CytoExploreR-Transformations.html


Let's try ARCSINH Transformation
```{r}
GFP_trans <- cyto_transformer_arcsinh(timepoint_gating_set,
                                      channels = c("B2-A"),
                                      widthBasis = -10
)#returns it as a list
FSC_SSC_trans <- cyto_transformer_arcsinh(timepoint_gating_set,
                                      channels = c("FSC-A", "FSC-H", "SSC-A", "SSC-H")
)
combined_trans <- cyto_transformer_combine(GFP_trans,FSC_SSC_trans)
transformed_timepoint_gating_set <- cyto_transform(timepoint_gating_set,
                                                   trans = combined_trans) #applies the the transformation and returns it as a gatingSet
```

```{r}
timepoint_raw_list <- cyto_extract(transformed_timepoint_gating_set, raw = TRUE, channels = c("FSC-A", "B2-A")) #raw flow data of each single cell as a list of matrices

arcsinh_norm = map_df(timepoint_raw_list, ~as.data.frame(.x), .id="name") %>% #convert to df, put list name in new column
  mutate(name = as.factor(name)) %>% #convert `name` to factor
  left_join(experiment_details %>% #join by name column to add metadata
  mutate(generation = as.factor(unique(experiment_details$generation)))) %>%
  mutate(B2A_FSC = `B2-A`/`FSC-A`)

arcsinh_norm%>%
  group_by(Description) %>%
  summarize(median(B2A_FSC))
```

Let's try log Transformation
```{r}
timepoint_gating_set <- cyto_setup(path = paste0(folders[1]), restrict=TRUE, select="fcs", details=F) #edit Markers on Viewer pane, Save & Close

experiment_details <- read_csv(exp_details_path) #import experiment-details.csv

for(i in 1:length(names(experiment_details))){
  flowWorkspace::pData(timepoint_gating_set)[names(experiment_details[i])]<-experiment_details[i]
}

GFP_trans <- cyto_transformer_log(timepoint_gating_set,
                                      channels = c("B2-A"),
                                      widthBasis = -100
)#returns it as a list
FSC_SSC_trans <- cyto_transformer_log(timepoint_gating_set,
                                      channels = c("FSC-A", "FSC-H", "SSC-A", "SSC-H")
)
combined_trans <- cyto_transformer_combine(GFP_trans,FSC_SSC_trans)
transformed_timepoint_gating_set <- cyto_transform(timepoint_gating_set,
                                                   trans = combined_trans) #applies the the transformation and returns it as a gatingSet

timepoint_raw_list <- cyto_extract(transformed_timepoint_gating_set, raw = TRUE, channels = c("FSC-A", "B2-A")) #raw flow data of each single cell as a list of matrices

log_norm = map_df(timepoint_raw_list, ~as.data.frame(.x), .id="name") %>% #convert to df, put list name in new column
  mutate(name = as.factor(name)) %>% #convert `name` to factor
  left_join(experiment_details %>% #join by name column to add metadata
  mutate(generation = as.factor(unique(experiment_details$generation)))) %>%
  mutate(B2A_FSC = `B2-A`/`FSC-A`)

log_norm%>%
  group_by(Description) %>%
  summarize(median(B2A_FSC))
```

Currently, we use logicle for B2-a and log for FSC,ssc
```{r}
timepoint_gating_set <- cyto_setup(path = paste0(folders[1]), restrict=TRUE, select="fcs", details=F) #edit Markers on Viewer pane, Save & Close

experiment_details <- read_csv(exp_details_path) #import experiment-details.csv

for(i in 1:length(names(experiment_details))){
  flowWorkspace::pData(timepoint_gating_set)[names(experiment_details[i])]<-experiment_details[i]
}

GFP_trans <- cyto_transformer_logicle(timepoint_gating_set,
                                      channels = c("B2-A"),
                                      widthBasis = -100
)#returns it as a list
FSC_SSC_trans <- cyto_transformer_log(timepoint_gating_set,
                                      channels = c("FSC-A", "FSC-H", "SSC-A", "SSC-H")
)
combined_trans <- cyto_transformer_combine(GFP_trans,FSC_SSC_trans)
transformed_timepoint_gating_set <- cyto_transform(timepoint_gating_set,
                                                   trans = combined_trans) #applies the the transformation and returns it as a gatingSet

timepoint_raw_list <- cyto_extract(transformed_timepoint_gating_set, raw = TRUE, channels = c("FSC-A", "B2-A")) #raw flow data of each single cell as a list of matrices

now = map_df(timepoint_raw_list, ~as.data.frame(.x), .id="name") %>% #convert to df, put list name in new column
  mutate(name = as.factor(name)) %>% #convert `name` to factor
  left_join(experiment_details %>% #join by name column to add metadata
  mutate(generation = as.factor(unique(experiment_details$generation)))) %>%
  mutate(B2A_FSC = `B2-A`/`FSC-A`)

now %>%
  group_by(Description) %>%
  summarize(median(B2A_FSC))
```




Try Arcsinh B2-A and Log FSC,SSC
```{r}
timepoint_gating_set <- cyto_setup(path = paste0(folders[1]), restrict=TRUE, select="fcs", details=F) #edit Markers on Viewer pane, Save & Close

experiment_details <- read_csv(exp_details_path) #import experiment-details.csv

for(i in 1:length(names(experiment_details))){
  flowWorkspace::pData(timepoint_gating_set)[names(experiment_details[i])]<-experiment_details[i]
}

GFP_trans <- cyto_transformer_arcsinh(timepoint_gating_set,
                                      channels = c("B2-A"),
                                      widthBasis = -100
)#returns it as a list
FSC_SSC_trans <- cyto_transformer_log(timepoint_gating_set,
                                      channels = c("FSC-A", "FSC-H", "SSC-A", "SSC-H")
)
combined_trans <- cyto_transformer_combine(GFP_trans,FSC_SSC_trans)
transformed_timepoint_gating_set <- cyto_transform(timepoint_gating_set,
                                                   trans = combined_trans) #applies the the transformation and returns it as a gatingSet

timepoint_raw_list <- cyto_extract(transformed_timepoint_gating_set, raw = TRUE, channels = c("FSC-A", "B2-A")) #raw flow data of each single cell as a list of matrices

arclog = map_df(timepoint_raw_list, ~as.data.frame(.x), .id="name") %>% #convert to df, put list name in new column
  mutate(name = as.factor(name)) %>% #convert `name` to factor
  left_join(experiment_details %>% #join by name column to add metadata
  mutate(generation = as.factor(unique(experiment_details$generation)))) %>%
  mutate(B2A_FSC = `B2-A`/`FSC-A`)

arclog %>%
  group_by(Description) %>%
  summarize(median(B2A_FSC))
```

#STEP 5:  Use function to perform analysis
#A function that will
#1 Read in all the files in a folder
#2 Read in experiment details files using pData
#3 Specify experiment markers
#4 Transform gating set
#5 Apply existing gating file using cyto_gatingTemplate_apply
#6.Output stats file as .csv
#Author: David & Julie

my_markers<-c("GFP") #list your marker name(s)
channel<-c("B2-A") #list your channel(s)
names(my_markers)<-channel

analyze_all_exp = function(folder_name, my_markers, gating_template="cytek_gating.csv") {

  path <- folder_name #gets relative path name for folder to be analyzed

  prefix <- folder_name %>% str_extract("([0-9])+_EE_GAP1_ArchMuts_2021") #extracts the time point number from folder name

  exp_details_path <- paste0(path,"/",prefix,"_experiment_details.csv") #gets experiment details .csv from correct directory
  #exp_details_path <- list.files(path = paste0(path), pattern = "_experiment_details.csv", full.names = T)
  #1. read in files and make a gating set
  print(path)
  timepoint_gating_set <- cyto_setup(path=path, select="fcs", details=F, markers = F)

  #2. read in experiment details for that gating set
  experiment_details <- read_csv(exp_details_path, show_col_types = F) #import experiment-details.csv
  #Write For Loop: for column in exp_details_path, add that column to timepoint_gating_set's metadata
  experiment_details <- read_csv(exp_details_path) #import experiment-details.csv
  for(i in 1:length(names(experiment_details))){
    flowWorkspace::pData(timepoint_gating_set)[names(experiment_details[i])]<-experiment_details[i]
  }
experiment_details
  #3. specify markers for that gating set
  markernames(timepoint_gating_set)<-my_markers

  #4. transform data
  GFP_trans <- cyto_transformer_logicle(timepoint_gating_set,
                                                 channels = c("B2-A"),
                                                 widthBasis = -10
                                                 )#returns it as a list
  FSC_SSC_trans <- cyto_transformer_log(timepoint_gating_set,
                        channels = c("FSC-A", "FSC-H", "SSC-A", "SSC-H")
                        )
  combined_trans <- cyto_transformer_combine(GFP_trans,FSC_SSC_trans)
  transformed_timepoint_gating_set <- cyto_transform(timepoint_gating_set,
                                       trans = combined_trans) #applies the the transformation and returns it as a gatingSet

  #5. apply gating-template.csv to transformed gating set
  cyto_gatingTemplate_apply(transformed_timepoint_gating_set, gatingTemplate= gating_template)
#  cyto_gatingTemplate_apply(transformed_timepoint_gating_set, gatingTemplate= "cytek_gating_01_02_04_v2.csv")

  #6. Get cell counts and frequencies inside each gate
  #Julie

  #get cell count from each gate
  gs_pop_get_stats(transformed_timepoint_gating_set, c("Single_cells", "zero_copy", "one_copy", "two_or_more_copy")) %>%
    rename(Gate = pop, name = sample, Count = count) %>%
    left_join(experiment_details) %>%
    write_csv(paste0(version_name,"_counts_",prefix,".csv"))

  #get frequency of cells inside each gate
  gs_pop_get_stats(transformed_timepoint_gating_set, c("Single_cells","zero_copy", "one_copy", "two_or_more_copy"), type = "percent") %>%
    rename(Gate = pop, name = sample, Frequency = percent) %>%
    left_join(experiment_details) %>%
    write_csv(paste0(version_name,"_freq_",prefix,".csv"))

  #get single cell fluorescence normalized over cell size
#  timepoint_raw_list <- cyto_extract(transformed_timepoint_gating_set, parent = "Single_cells", raw = TRUE, channels = c("FSC-A", "B2-A")) #raw flow data of each single cell as a list of matrices

#  map_df(timepoint_raw_list, ~as.data.frame(.x), .id="name") %>% #convert to df, put list name in new column
#    mutate(name = as.factor(name)) %>% #convert `name` to factor
#    left_join(experiment_details %>% #join by name column to add metadata
#                mutate(generation = as.factor(unique(experiment_details$generation)))) %>%
#    mutate(B2A_FSC = `B2-A`/`FSC-A`) %>% #compute normalized fluor
#    write_csv(paste0(version_name,"_SingleCellDistributions_",prefix,".csv"))


}

#STEP 6:  Apply function from STEP 5 to all subdirectories
#Uses map from purr() to apply function from step 5 to all directories
#Author: Julie

try(map(folders[2:length(folders)],analyze_all_exp, my_markers, gating_template = paste0("cytek_gating_",version_name,".csv")))

#STEP 7: Pull in all counts or freq or single cell distribution files from directory and combine into a single dataframe
#Author: Julie
#01_02_04_v2_fw_counts_all_timepoints
#01_02_04_v2_fw_freq_all_timepoints
list.files(path = ".", pattern = paste0(version_name,"_counts_([0-9])+_EE_GAP1_ArchMuts_2021")) %>%
  read_csv() %>%
  mutate(gating_template = paste0("cytek_gating_",version_name,".csv")) %>%
  write_csv(file = paste0(version_name,"_counts_all_timepoints.csv")) #write a function to establish/customize the version file name then can paste0(version_filename)

list.files(path = ".", pattern = paste0(version_name,"_freq_([0-9])+_EE_GAP1_ArchMuts_2021")) %>%
  read_csv() %>%
  mutate(gating_template = paste0("cytek_gating_",version_name,".csv")) %>%
  write_csv(file = paste0(version_name,"_freq_all_timepoints.csv"))

## Do on hpc because large files, do once. Don't even try to run this command on your laptop. 12GB file.
# list.files(path = ".", pattern = paste0(version_name,"_SingleCellDistributions")) %>%
#   read_csv() %>%
#   mutate(gating_template = paste0("cytek_gating_",version_name,".csv")) %>%
#   write_csv(file = paste0(version_name,"_SingleCellDistributions_all_timepoints.csv"))


```{r}
?data.frame(colname = vctor1, colname2 = vector2)
```

